---
title: "Linear Regression"
arren Wuffett is seeking new investment opportunities. He collected some data from start-ups and wants to know how their business strategies (spendings
in marketing, administration, and research and development) influence their
profit. Help him find a model to predict profit by using a simple linear regression and the mentioned variables. Hint: in R use the commands lm, predict,
cor.
---

---
A Download and parse the corresponding data set ex03 startups.csv from
Moodle. Explore the pairwise correlations in the data and build 4 linear
models in which you use the variables Marketing.Spend, Administration,
R.D.Spend or all of them to predict profit
---


```{r}
startups_data <- read.csv("D:\\data\\Computational statistics\\startups.csv", sep = ",")
head(startups_data)
startups_data <- subset(startups_data, select = -c(State))
cor(startups_data)

pairs(startups_data)


lm_R.D.Spend <- lm(Profit~R.D.Spend, data = startups_data)
lm_Administration <- lm(Profit~Administration, data = startups_data)
lm_Marketing.Spend <- lm(Profit~Marketing.Spend, data = startups_data)
lm_all <- lm(Profit~., data = startups_data)
```


---
B) Examine your prediction results visually and by computing the correlation coefficients between the predicted and observed data. Which model
performs best?1 Based on the best model, what is the increase in profit
when you increase the investments in R.D. by $100?
---

```{r}
#Examine by visualization

library ( ggplot2 )
require ( gridExtra )

p1 <- ggplot(startups_data, aes(x=predict(lm_R.D.Spend, startups_data), y=Profit))+
  geom_point()+geom_smooth(method = lm, se=FALSE)+
  xlab("R.D.Spend") + ylab("Profit") +ggtitle("lm of R.D.Spend")

p2 <- ggplot(startups_data, aes(x=predict(lm_Administration, startups_data), y=Profit))+
  geom_point()+geom_smooth(method = lm, se=FALSE)+
  xlab("Administration") + ylab("Profit") +ggtitle("lm of Administration")

p3 <- ggplot(startups_data, aes(x=predict(lm_Marketing.Spend, startups_data), y=Profit))+
  geom_point()+geom_smooth(method = lm, se=FALSE)+
  xlab("Marketing.Spend") + ylab("Profit") +ggtitle("lm of Marketing Spend")

p4 <- ggplot(startups_data, aes(x=predict(lm_all, startups_data), y=Profit))+
  geom_point()+geom_smooth(method = lm, se=FALSE)+
  xlab("ALL") + ylab("Profit") +ggtitle("lm of All Variables")

grid.arrange(p1, p2, p3, p4, ncol=2)


```
---
based on the linear regression model R.D.Spend is highly correleted with profit, so we can invest more on R.D.Spend

---
```{r}
#predicting correlation with correlation coefficient

predictions <- data.frame("profit" = startups_data$Profit,
                          "rd_pred" = predict(lm_R.D.Spend, startups_data),
                          "Administration" = predict(lm_Administration, startups_data),
                          "Marketing Spend" = predict(lm_Marketing.Spend, startups_data),
                          "ALL" = predict(lm_all, startups_data))

cor(predictions)
```

---
The model based on all explanatory variables performs best, although it gives
only a minor improvement over the prediction results based on the R.D. Spend.
Based on this information we can now increase the investment in R.D. by $100
for this model.
---
```{r}
lm_all$coefficients["R.D.Spend"]*100
```
---
by spending 100 dollars on R.D.Spend, we expect 80 dollar profit in overall.
---


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
